{ \sffamily \LARGE Electronic supplementary material}\\{\bf \sffamily \LARGE Evolution of reciprocity with limited payoff memory

### 2023-10-26 paper/si.tex

\documentclass[11pt]{article}
\usepackage[dvipsnames]{xcolor}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,array,eurosym,multirow}
\usepackage[sort&compress,numbers]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[margin=2.5cm, includefoot, footskip=30pt]{geometry}
\usepackage{standalone}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{blkarray}
\usepackage[ruled,vlined]{algorithm2e}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}
\newcommand{\R}{\mathbb{R}}

@@

\def\alld{\texttt{ALLD}}
\def\tft{\texttt{TFT}}
\def\gtft{\texttt{GTFT}}
\def\rolemodel{\text{R}}
\def\learner{\text{L}}
\def\resident{\texttt{R}}
\def\mutant{\texttt{M}}
\def\strategy{\texttt{$\sigma$}}
\def\esm{electronic supplementary material}

@@

\usepackage{tikz}
\usetikzlibrary{arrows}

@@

\newtheoremstyle{plainCl1}% name
{9pt}%      Space above, empty = 'usual value'
{9pt}%      Space below
{\it}% 	   Body font
{}%         Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}% Thm head font
{.}%        Punctuation after thm head
{0.2cm}% Space after thm head: \newline = linebreak
{}%         Thm head spec

@@

\noindent
This document provides further details on our methods and derivations, and it contains additional simulation results. 
Section~\ref{section:model} summarizes the model. 
In particular, we provide further details on our implementation of the evolutionary dynamics, and our use of the rare-mutations limit. 
In Section~\ref{section:analyticalresults} we derive analytical results for the various settings we consider.
These settings differ in what kind of payoff information individuals take into account when updating their strategies. 
In the perfect memory setting, individuals take into account all their interactions against all co-players.
In the limited memory setting, they only consider the very last round of their very last interaction.  
In addition, we describe several model extensions in which the amount of information taken into account is in between these two extremes. 
Finally, Section~\ref{section:furthersimulations} presents further simulation results. 
In particular, we confirm that our main results continue to hold ({\it i}) when players use memory-1 strategies instead of reactive strategies, and  ({\it ii}) when mutations are no longer rare.

@@

\noindent
{\bf Summary of the model.} As described in the main text, we study cooperative behavior in a population of size $N$.
The dynamics unfolds on two time-scales. 
The short time scale describes the game dynamics. 
Here individuals are randomly matched to interact in a repeated prisoner's dilemma.  
Each round individuals can choose whether to cooperate ($C$) or defect ($D$). 
In the most general setting, the resulting one-shot payoffs can be summarized by the payoff matrix 
\begin{equation}\label{eq:prisoner}
    \begin{blockarray}{ccc}
        & \text{cooperate} & \text{defect} \\
        \begin{block}{c(cc)}
            \text{cooperate} & R & S \\
            \text{defect} & T & P \\
        \end{block}
    \end{blockarray}.
\end{equation}
Here, $R$ is interpreted as the reward for mutual cooperation, $S$ is the sucker's payoff, $T$ is the temptation, and $P$ is the punishment payoff~\citep{axelrod1981evolution}. 
Throughout this work, we parametrize these payoffs as $R\!=\!b\!-\!c$, $S\!=\!-c$, $T\!=\!b$, and $P\!=\!0$, where $b$ and $c$ are the benefit and cost of cooperation, respectively, with $b\!>\!c\!>0$. 
After each round, players learn their co-player's previous action. 
Then the game continues for another round with probability~$\delta$.
Players make their decisions whether to cooperate in any given round based on their reactive strategies $\mathbf{s}\!=\!(y,p,q)$. 
The entry $y$ determines a player's first-round cooperation probability. 
The other entries $p$ and $q$ determine the player's cooperation probability in all subsequent rounds. 
The probability is $p$ if the co-player cooperated in the previous round, and it is $q$ if the co-player defected. 
On the short time scale, the players' strategies are fixed.

@@

{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
The case of perfect payoff memory corresponds to the classical case considered in the previous literature. 
Here, individuals update their strategies based on the expected payoffs, based on all rounds and all possible interaction partners. 
When players use reactive strategies (or more generally, strategies with finite memory), these expected payoffs can be computed explicitly, based on a Markov chain approach~\cite{sigmund2010calculus}. 
To this end, consider two players with strategies $\strategy_1\!=\!(y_1, p_1, q_1$) and $\strategy_2\!=\!(y_2,p_2,q_2)$, respectively. 
In each round $t$ of the game, player~1 may get one of the four possible payoffs $R$, $S$, $T$, or $P$, as described by the general payoff matrix~\eqref{eq:prisoner}. 
Let $\mathbf{v}(t)\!=\!\big(v_R(t),v_S(t),v_T(t),v_P(t)\big)$ denote the respective probability distribution of observing one of these four outcomes.  
This probability distribution can be computed recursively. 
Using the shortcut notation $\bar{z}\!=\!1\!-\!z$ for any $z\!\in\![0,1]$, we get for the initial round 
\begin{equation}
\mathbf{v}_0:=\mathbf{v}(0) = (y_1 y_2,~y_1 \bar{y}_2,~\bar{y}_1 y_2,~\bar{y}_1 \bar{y}_2).
\end{equation}
Given $\mathbf{v}(t)$, we can compute $\mathbf{v}(t\!+\!1)$ as  
\begin{equation}
\mathbf{v}(t\!+\!1) = \mathbf{v}(t)\cdot M,
\end{equation}
where $M$ is the transition matrix of the process, 
\begin{equation}\label{eq:transition_matrix}
  \input{static/matrix.tex}.
\end{equation}
Based on this recursion, we can compute how often player~1 receives one of the four payoffs $R$, $S$, $T$, $P$ on average (across all possible realizations of games among the two players). This average distribution $\mathbf{v}$ is
\begin{equation}
\mathbf{v} := (1\!-\!\delta) \sum_{t=0}^\infty \delta^t \mathbf{v}(t) 
= (1\!-\!\delta) \mathbf{v_0} \sum_{t=0}^\infty \delta^t M^t
= (1\!-\!\delta) \mathbf{v_0} (I_4-\delta M)^{-1},
\end{equation}
where $I_4$ is the $4\times4$ identity matrix. 
Based on this general formula, the four entries of $\mathbf{v}=(v_R, v_S, v_T, v_P)$ can be computed explicitly. 
Using the shortcut notation $r_i:=p_i\!-\!q_i$, we obtain
    \begin{equation} \label{Eq:vAverage}
      \setlength{\arraycolsep}{1pt}
      \begin{array}{rcl}    
      v_{R} &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
      v_{S}&= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
      v_{T} &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
      v_{P} &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
      \end{array}
    \end{equation}
Using this distribution $\mathbf{v}$, we compute player 1's expected payoff as the weighted average
\begin{equation}
\pi(\strategy_1,\strategy_2) = R \, v_R + S\,v_S + T\,v_T + P\,v_P. 
\end{equation}

@@

~\\
{\bf Invasion Analysis.}
As an application of this formalism, we can compute when cooperation is stochastically stable in the perfect information setting. 
To this end, suppose there is only a single mutant, $k\!=\!1$. The residents adopt Generous Tit-for-Tat, $\gtft = (1, 1, q)$ and the mutant adopts $\alld\! = \! (0,0,0)$. When two GTFT players interact, the resulting average distribution according to Eq.~\eqref{Eq:vAverage} simplifies to
\begin{align*}
\mathbf{v}(\gtft,\gtft) = (1, 0, 0, 0).
\end{align*}
On the other hand, if \alld{} interacts with \gtft, the respective probabilities become,
\begin{align*}
  \mathbf{v}(\alld,\gtft) = \big(0,~0,~1\!-\!\delta\!+\!\delta q,~\delta(1\! -\! q)\big).
\end{align*}
Based on Eq.~\eqref{Eq:ExpPay}, we can compute the strategies' expected payoffs as
\begin{align*}
\pi_\gtft =  \frac{N\!-\!2}{N-1} (b - c)  \!-\! \frac{1}{N-1}\big(1\!-\!\delta\!+\!\delta q\big)c
\quad \text{and} \quad
\pi_\alld  = \big(1\!-\!\delta\!+\!\delta q\big)b.
\end{align*}
As a consequence, we can calculate the corresponding ratio of transition probabilities according to Eq.~\eqref{eq:LambdaRatio},
\begin{align*}
\frac{\lambda_1^{-}}{\lambda_1^{+}} = e^{-\beta \left( (1-\delta+\delta q)(b+\frac{c}{N-1})- \frac{N-2}{N-1}(b-c) \right)}
\end{align*}
By definition, cooperation is stochastically stable if this ratio exceeds one, which is equivalent to
\begin{equation*}
q<1-\frac{1}{\delta}\cdot \frac{b+(N\!-\!1)c}{(N\!-\!1)b+c}.
\end{equation*}
For such a strategy to be feasible we require $q\!>\!0$, which implies $\delta \!>\! \big(b+(N\!-\!1)c\big)/\big((N\!-\!1)b\!+\!c\big)$.
In particular, in the limit of large populations \(N \!\rightarrow\! \infty\), we obtain that cooperation is stochastically stable if  \(q \!<\! 1\! -\! c/(\delta b)\). 
The minimum continuation probability for such a strategy to exist is $\delta \!>\! c/b$. 
In this way, we recover the classical conditions for cooperation to be feasible under direct reciprocity~\citep{molander:jcr:1985,Nowak1992tit,Schmid:NHB:2021}.

@@

In the expected payoffs case the payoff of a pair depends on the average payoff
they received over an infinite number of turns. In the limited payoff memory,
the payoff of a pair depends on the average payoffs they received in the last
turn. Moreover, in expected payoffs it is assumed that a player interacts with
every member of the population whereas in the limited payoff memory approach a
player has one interaction.

Initially, we define the probability that a reactive strategy receives the
payoff $u\!\in\! \mathcal{U}$ in the very last round of the game against another
reactive strategy (Proposition~\ref{proposition:last_round}).

\begin{Prop}\label{proposition:last_round} Consider a repeated game, with
    continuation probability $\delta$, between players with reactive strategies
    $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
    the probability that the $s_1$ player receives the payoff $u\!\in\!
    \mathcal{U}$ in the very last round of the game is given by
    $v_{u}(s_1,s_2)$, as given by Eq.~(\ref{Eq:LastRound}).

    \begin{equation} \label{Eq:LastRound}
      \setlength{\arraycolsep}{1pt}
      \begin{array}{rcl}

      v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]

      v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]

      v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]

      v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
      \end{array}
    \end{equation}

In these expressions, we have used the notation $l_i:=p_i\!-\!q_i$,
$\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
$\bar{l}_i:=\bar{p}_i\!-\!\bar{q}_i=-l_i$ for $i\!\in\!\{1,2\}$.
\end{Prop}

Note that in the proposition we here we focus on the case of the donation game/prisoner's
dilemma but the result applies to any \(2 \times 2\) symmetric game.

\begin{proof}
Given a play between two reactive strategies with continuation probability
$\delta$. The outcome at turn \(t\) is given by,

\begin{equation}\label{eq:}
  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)},
\end{equation}

where $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
the very first round, and \(1- \delta\) the probability that the game ends. It
can be shown that,

\begin{align*}
  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)} & = (1 - \delta)(\mathbf{v_0} + \delta \mathbf{v_0} M + \delta^{2}\mathbf{v_0} M ^{2} + \dots )\\ 
   & = (1 - \delta)\mathbf{v_0} (1 + \delta M + \delta^{2}M ^{2} + \dots ) \text{ using standard formula for geometric series}\\ 
   & = (1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}
\end{align*}

where \((1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}\) is vector \(\in R^{4}\)
and it the probabilities for being in any of the outcomes \(CC, CD, DC, DD\) in
the last round. Combining this with the payoff vector \(u\) and some algebraic
manipulation we derive to the Equation~\ref{Eq:LastRound}.
\end{proof}

>>>>

\documentclass[11pt]{article}
\usepackage[dvipsnames]{xcolor}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,array,eurosym,multirow}
\usepackage[sort&compress,numbers]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[margin=2.5cm, includefoot, footskip=30pt]{geometry}
\usepackage{standalone}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{blkarray}
\usepackage[ruled,vlined]{algorithm2e}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}
\newcommand{\R}{\mathbb{R}}
\usepackage{tikz}
\usetikzlibrary{arrows}

@@

\def\alld{\texttt{ALLD}}
\def\tft{\texttt{TFT}}
\def\gtft{\texttt{GTFT}}
\def\rolemodel{\text{RM}}
\def\learner{\text{L}}
\def\resident{\texttt{R}}
\def\mutant{\texttt{M}}
\def\strategy{s}
\def\esm{electronic supplementary material}

@@

\renewcommand{\figurename}{Supplementary Figure}

@@

\newtheoremstyle{plainCl1}% name
{9pt}%      Space above, empty = 'usual value'
{9pt}%      Space below
{}% 	   Body font
{}%         Indent amount (empty = no indent, \parindent = para indent)
{\it}% Thm head font
{.}%        Punctuation after thm head
{0.2cm}% Space after thm head: \newline = linebreak
{}%         Thm head spec

@@

\noindent
This document provides further details on our methods and derivations, and it contains additional simulation results. 
Section~\ref{section:model} summarizes the model. 
In particular, we provide further details on our implementation of the evolutionary dynamics, and our use of the rare-mutations limit. 
In Section~\ref{section:analyticalresults} we derive analytical results for the various settings we consider.
These settings differ in what kind of payoff information individuals take into account when updating their strategies. 
In the perfect memory setting, individuals take into account all their interactions against all co-players.
In the limited memory setting, they only consider the very last round of their very last interaction.  
In addition, we describe several model extensions in which the amount of information taken into account is in between these two extremes. 
Finally, Section~\ref{section:furthersimulations} presents further simulation results. 
In particular, we confirm that our main results continue to hold ({\it i}) when mutations are no longer rare, and  ({\it ii}) when players use memory-1 strategies instead of reactive strategies.

@@

\noindent
{\bf Summary of the model.} As described in the main text, we study cooperative behavior in a population of size $N$, with $N$ being even.
The dynamics unfolds on two time-scales. 
The short time scale describes the game dynamics. 
Here the $N$ individuals are randomly matched to form $N/2$ pairs to interact in a repeated prisoner's dilemma.  
Each round, individuals can choose whether to cooperate ($C$) or defect ($D$). 
In the most general setting, the resulting one-shot payoffs can be summarized by the payoff matrix 
\begin{equation}\label{eq:prisoner}
    \begin{blockarray}{ccc}
        & \text{cooperate} & \text{defect} \\
        \begin{block}{c(cc)}
            \text{cooperate} & R & S \\
            \text{defect} & T & P \\
        \end{block}
    \end{blockarray}.
\end{equation}
Here, $R$ is interpreted as the reward for mutual cooperation, $S$ is the sucker's payoff, $T$ is the temptation, and $P$ is the punishment payoff~\citep{axelrod1981evolution}. 
Throughout this work, we parametrize these payoffs as $R\!=\!b\!-\!c$, $S\!=\!-c$, $T\!=\!b$, and $P\!=\!0$, where $b$ and $c$ are the benefit and cost of cooperation, respectively, with $b\!>\!c\!>0$. 
After each round, players learn their co-player's previous action. 
Then the game continues for another round with probability~$\delta$.
Players make their decisions whether to cooperate in any given round based on their reactive strategies $\mathbf{s}\!=\!(y,p,q)$. 
The entry $y$ determines a player's first-round cooperation probability. 
The other entries $p$ and $q$ determine the player's cooperation probability in all subsequent rounds. 
The probability is $p$ if the co-player cooperated in the previous round, and it is $q$ if the co-player defected. 
On this short time scale, the players' strategies are fixed, and players are consecutively matched to play many repeated games with randomly changing interaction partners.

@@

{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
The case of perfect payoff memory corresponds to the classical case considered in the previous literature. 
Here, individuals update their strategies based on the expected payoffs, based on all rounds and all possible interaction partners. 
When players use reactive strategies (or more generally, strategies with finite memory), these expected payoffs can be computed explicitly, based on a Markov chain approach~\cite{sigmund2010calculus}. 
To this end, consider two players with strategies $\strategy_1\!=\!(y_1, p_1, q_1$) and $\strategy_2\!=\!(y_2,p_2,q_2)$, respectively. 
In each round $t$ of the game, player~1 may get one of the four possible payoffs $R$, $S$, $T$, or $P$, as described by the general payoff matrix~\eqref{eq:prisoner}. 
Let $\mathbf{v}(t)\!=\!\big(v_R(t),v_S(t),v_T(t),v_P(t)\big)$ denote the respective probability distribution of observing one of these four outcomes.  
This probability distribution can be computed recursively. 
Using the shortcut notation $\bar{z}\!=\!1\!-\!z$ for any $z\!\in\![0,1]$, we get for the initial round 
\begin{equation} \label{Eq:v0}
\mathbf{v}_0:=\mathbf{v}(0) = (y_1 y_2,~y_1 \bar{y}_2,~\bar{y}_1 y_2,~\bar{y}_1 \bar{y}_2).
\end{equation}
Given $\mathbf{v}(t)$, we can compute $\mathbf{v}(t\!+\!1)$ as  
\begin{equation} \label{Eq:vt}
\mathbf{v}(t\!+\!1) = \mathbf{v}(t)\cdot M,
\end{equation}
where $M$ is the transition matrix of the process, 
\begin{equation} \label{eq:transition_matrix}
M=\left[
\begin{array}{llll}
p_1 p_2	&p_1 \bar{p}_2	&\bar{p}_1 p_2	&\bar{p}_1\bar{p}_2\\
q_1 p_2	&q_1 \bar{p}_2 &\bar{q}_1 p_2	&\bar{q}_1\bar{p}_2\\
p_1 q_2	&p_1 \bar{q}_2	&\bar{p}_1 q_2	&\bar{p}_1\bar{q}_2\\
q_1 q_2	&q_1 \bar{q}_2	&\bar{q}_1 q_2	&\bar{q}_1\bar{q}_2
\end{array}
\right]
\end{equation}
Based on this recursion, we can compute how often player~1 receives one of the four payoffs $R$, $S$, $T$, $P$ on average (across all possible realizations of games among the two players). This average distribution $\mathbf{v}$ is
\begin{equation} \label{Eq:vDefinition}
\mathbf{v} := (1\!-\!\delta) \sum_{t=0}^\infty \delta^t \mathbf{v}(t) 
= (1\!-\!\delta) \mathbf{v_0} \sum_{t=0}^\infty \delta^t M^t
= (1\!-\!\delta) \mathbf{v_0} (I_4-\delta M)^{-1},
\end{equation}
where $I_4$ is the $4\times4$ identity matrix. 
Based on this general formula, the four entries of $\mathbf{v}=(v_R, v_S, v_T, v_P)$ can be computed explicitly. 
Using the shortcut notation $r_i:=p_i\!-\!q_i$, we obtain
    \begin{equation} \label{Eq:vAverage}
      \setlength{\arraycolsep}{1pt}
      \begin{array}{rcl}    
      v_{R} &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
      v_{S}&= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2-r_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
      v_{T} &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1-r_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
      v_{P} &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1-r_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2-r_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
      \end{array}
    \end{equation}
Using this distribution $\mathbf{v}$, we compute player 1's expected payoff as the weighted average
\begin{equation}
\pi(\strategy_1,\strategy_2) = R \, v_R + S\,v_S + T\,v_T + P\,v_P. 
\end{equation}

@@

~\\
{\bf Stochastic stability of cooperation.}
As an application of this formalism, we can compute when cooperation is stochastically stable in the perfect information setting. 
To this end, suppose there is only a single mutant, $k\!=\!1$. The residents adopt Generous Tit-for-Tat, $\gtft = (1, 1, q)$ and the mutant adopts $\alld\! = \! (0,0,0)$. When two GTFT players interact, the resulting average distribution according to Eq.~\eqref{Eq:vAverage} simplifies to
\begin{align*}
\mathbf{v}(\gtft,\gtft) = (1, 0, 0, 0).
\end{align*}
On the other hand, if \alld{} interacts with \gtft, the respective probabilities become,
\begin{align*}
  \mathbf{v}(\alld,\gtft) = \big(0,~0,~1\!-\!\delta\!+\!\delta q,~\delta(1\! -\! q)\big).
\end{align*}
Based on Eq.~\eqref{Eq:ExpPay}, we can compute the strategies' expected payoffs as
\begin{align*}
\pi_\gtft =  \frac{N\!-\!2}{N-1} (b - c)  \!-\! \frac{1}{N-1}\big(1\!-\!\delta\!+\!\delta q\big)c
\quad \text{and} \quad
\pi_\alld  = \big(1\!-\!\delta\!+\!\delta q\big)b.
\end{align*}
As a consequence, we can calculate the corresponding ratio of transition probabilities according to Eq.~\eqref{eq:LambdaRatio},
\begin{align*}
\frac{\lambda_1^{-}}{\lambda_1^{+}} = e^{-\beta \left( (1-\delta+\delta q)(b+\frac{c}{N-1})- \frac{N-2}{N-1}(b-c) \right)}
\end{align*}
By definition, cooperation is stochastically stable if this ratio exceeds one, which is equivalent to
\begin{equation*}
q<1-\frac{1}{\delta}\cdot \frac{b+(N\!-\!1)c}{(N\!-\!1)b+c}.
\end{equation*}
For such a strategy to be feasible we require $q\!>\!0$, which implies $\delta \!>\! \big(b+(N\!-\!1)c\big)/\big((N\!-\!1)b\!+\!c\big)$.
In particular, in the limit of large populations \(N \!\rightarrow\! \infty\), we obtain that cooperation is stochastically stable if  \(q \!<\! 1\! -\! c/(\delta b)\). 
The minimum continuation probability for such a strategy to exist is $\delta \!>\! c/b$. 
In this way, we recover the classical conditions for cooperation to be feasible under direct reciprocity~\citep{molander:jcr:1985,Nowak1992tit,Schmid:NHB:2021}.

@@

%% RESULTS: Limited memory:  problem formulation %%

{\bf Computing the distribution of last-round payoffs.}
The case of perfect payoff memory is straightforward to handle; here, every player gets the expected payoff with certainty. 
In comparison, computing transition probabilities for the case of limited payoff memory is more elaborate. 
Here, we need to consider the different possible outcomes that both the learner and role model may have experienced in their very last interaction. 
A further complication arises when the learner's last interaction partner happens to be the role model. 
In that case, the learner's and the role model's last payoff will be correlated (e.g., if the learner got the sucker's payoff of $S$, the role model's payoff is $T$ with certainty). 
To treat the case of limited memory analytically, let $\mathcal{U} \!=\! \{R,S,T,P\}$ be the set of possible one-shot payoffs. 
We first  compute the probability that a player receives a payoff of $u\!\in\!\mathcal{U}$ in the final round of the repeated game when the strategies of the two players are given.

%% RESULTS: Limited memory: probability distribution for final payoff for two fixed strategies%%

\begin{Prop}\label{proposition:last_round} 
Consider a repeated game with continuation probability $\delta$, between two players with reactive strategies $\strategy_1\!=\!(y_1, p_1, q_1$)  and $\strategy_2\!=\!(y_2,p_2,q_2)$. 
Let $\tilde{v}_u\!=\!\tilde{v}_u(\strategy_1,\strategy_2)$ denote the probability that the first player receives the payoff $u\!\in\!\mathcal{U}$ in the final round of the game. 
Then this probability  $\tilde{v}_u$ is identical to the player's average probability $v_u$ to receive payoff $u$ across all rounds of the game, as specified by Eq.~\eqref{Eq:vAverage}.
\end{Prop}

\begin{proof}
Let $\mathbf{\tilde{v}}:=(\tilde{v}_R, \tilde{v}_S, \tilde{v}_T, \tilde{v}_P)$. 
To compute $\mathbf{\tilde{v}}$, we first compute the probability to receive each outcome $u\!\in\!\mathcal{U}$ given that the last round happens to be round $\tau\!\in\!\{0,1,\ldots\}$. 
In that case, it follows from Eqs.~\eqref{Eq:v0} and~\eqref{Eq:vt} that the respective probability is
\begin{equation} \label{Eq:vTildeSimple}
\mathbf{\tilde v}(\tau) = \mathbf{v_0}\cdot M^\tau.
\end{equation}
Round $\tau$ is the last round with probability $\delta^\tau(1\!-\!\delta)$. 
By the law of total probability, we can obtain $\mathbf{\tilde{v}}$ by summing up over all possible last rounds,
\begin{equation}
\tilde{v} = \sum_{\tau=0}^\infty \delta^\tau(1\!-\!\delta)  \cdot \mathbf{v_0}M^\tau 
=  (1\!-\!\delta)\mathbf{v_0} \sum_{t=0}^\infty \delta^t  M^t. 
\end{equation}
This expression for $\mathbf{\tilde{v}}$ coincides with the expression for $\mathbf{v}$ in Eq.~\eqref{Eq:vDefinition}.
\end{proof}

\noindent
Note that while we focus on the prisoner's dilemma in this work, the Proposition applies to any repeated matrix game.\\

%% RESULTS: Limited memory - probability distribution in population setup %%

\noindent
{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
After these preparations, let us again consider the corresponding population setup, with $N\!-\!k$ residents with strategy  \(\strategy_\resident = (y_\resident, p_\resident, q_\resident)\) and $k$ mutants with strategy  \(\strategy_\mutant = (y_\mutant, p_\resident, q_\mutant)\).
At each step of the evolutionary process we choose a learner and a role model. 
The learner compares the performance of its strategy by comparing its own last one-shot payoff with the last one-shot payoff of the role model. 
In the following, we assume either the learner or the role model is a resident and that the other player is a mutant (otherwise it is certain that the number of mutants does not change). 
There are two major cases to consider.

%% RESULTS: Limited payoff memory - Different cases to consider %%

\begin{enumerate}
\item The learner and the role model have their last respective interaction with each other. This happens with probability $1/(N\!-\!1)$. In that case, there are four possible cases for their joint final payoffs, $(u_\resident,u_\mutant)\!\in\!\mathcal{U}_F\!:=\!\big\{(R,R),\,(S,T),\,(T,S)\,(P,P)\big\}$. These outcomes are distributed according to the distribution $\mathbf{\tilde{v}}(\strategy_\resident,\strategy_\mutant)$, as described in Proposition~\ref{proposition:last_round}.  
\item The learner's last interaction was not with the role model, with probability $(N\!-\!2)/(N\!-\!1)$. In this case, there are four different subcases, depending on whether the resident's last interaction partner was a mutant or a resident, and depending on whether the mutant's last interaction partner was a mutant or a resident. As a result, the resident's last one-shot payoff is distributed according to $\mathbf{\tilde{v}}(\strategy_\resident,\strategy_\mutant)$ or $\mathbf{\tilde{v}}(\strategy_\resident,\strategy_\resident)$; the mutant's last payoff is distributed according to $\mathbf{\tilde{v}}(\strategy_\mutant,\strategy_\mutant)$ or $\mathbf{\tilde{v}}(\strategy_\mutant,\strategy_\resident)$, respectively. 
\end{enumerate}
